---
title: "DSA 2021 Checks"
author: "Assil BEN-AMOR"
date : December 23, 2021
output: 
  html_document:
    code_folding: "none"
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
rm(list = ls());cat('\014')

if (!require(pacman)) install.packages(pacman)

pacman::p_load(readxl,
              rio,
              tidyverse,
              ggplot2)

options(scipen = 999)

options(dplyr.summarise.inform = FALSE)

file.names = list()

# file.names$data_1 = "input/data/DSA2021_Galmudug_CleanData.xlsx"
# file.names$data_2 = "input/data/DSA2021_Puntland_CleanData.xlsx"
# file.names$data_3 = "input/data/DSA2021_Somaliland_CleanData.xlsx"
file.names$tool = "input/tool/REACH_SOM_DSA_Survey_Tool_v7.xlsx"


file.names$data_1 = "/Users/mac/One drive/ACTED/Ahmed MAHAMOUD - 8_HQ-Validation/DSA2021_Galmudug_CleanData.xlsx"
file.names$data_2 = "/Users/mac/One drive/ACTED/Ahmed MAHAMOUD - 8_HQ-Validation/DSA2021_Puntland_CleanData.xlsx"
file.names$data_3 = "/Users/mac/One drive/ACTED/Ahmed MAHAMOUD - 8_HQ-Validation/DSA2021_Somaliland_CleanData.xlsx"

data_Galmudug <-  read_xlsx(file.names$data_1,
                            guess_max = 50000,
                            na = c("NA","#N/A",""," "),
                            sheet = "Clean Data")

data_Puntland <-  read_xlsx(file.names$data_2,
                            guess_max = 50000,
                            na = c("NA","#N/A",""," "),
                            sheet = "Clean Data")

data_Somaliland <-  read_xlsx(file.names$data_3,
                            guess_max = 50000,
                            na = c("NA","#N/A",""," "),
                            sheet = "Clean Data")

# names(data_Galmudug) <- gsub("^X_","_",names(data_Galmudug))
# names(data_Puntland) <- gsub("^X_","_",names(data_Puntland))
names(data_Puntland) <- gsub("/",".",names(data_Puntland))
names(data_Somaliland) <- gsub("/",".",names(data_Somaliland))


names(data_Galmudug) [! names(data_Galmudug)  %in% names(data_Somaliland)]
names(data_Somaliland) [! names(data_Somaliland)  %in% names(data_Galmudug)]


data <- do.call("rbind",list(data_Galmudug,data_Puntland,data_Somaliland)) %>% setNames(
   gsub("/",".",colnames(.))
)

names(data) <- gsub("^(X_|_)","",names(data))


questions <- read_excel(file.names$tool, sheet="survey", col_types="text") %>% 
  filter(!is.na(name)) %>% 
  mutate(q.type=as.character(lapply(type, function(x) str_split(x, " ")[[1]][1])),
         list_name=as.character(lapply(type, function(x) str_split(x, " ")[[1]][2])),
         list_name=ifelse(str_starts(type, "select_"), list_name, NA))

choices <- read_excel(file.names$tool, sheet="choices", col_types="text") %>% 
  filter(!is.na(list_name)) %>% select(c(list_name, name, "label::English (en)")) %>% 
  distinct()

listOfNumericQuestions <- questions %>% filter(type %in% c("integer","numeric","calculate")) %>% pull(name)

numericQuestions_index <- which(colnames(data) %in% listOfNumericQuestions)
data[numericQuestions_index] <- sapply(data[numericQuestions_index],as.numeric)

multipleSelectQuestions <- grep(pattern="\\.", colnames(data))
data[multipleSelectQuestions] <- sapply(data[multipleSelectQuestions],as.numeric)


generate_from_binaries <- function(data,select_multiple_questions) {
  
  do.call("cbind",map(select_multiple_questions,function(question_name,data) {
    df <- data %>% select(grep(paste0("^",question_name,"\\."), names(data))) 
    colnames(df) <- gsub(paste0(question_name,"\\."),"",colnames(df))
    map2_df(df, names(df), ~  replace(.x, .x==1, .y) %>% replace(. == 0, NA)) %>% 
      unite(!!sym(question_name),names(df),remove=TRUE,na.rm=TRUE,sep=" ") %>% 
      as.data.frame() %>% 
      mutate_all(list(~na_if(.,""))) 
  },data)) 
  
}

data <- sjmisc::replace_columns(data, 
                        generate_from_binaries(data,
                                               questions %>% filter(grepl("^select_multiple",type)) %>%
                                                 pull(name))) 

colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s; font-size:18.0pt;'>%s</span>", color,
            x)
  } else x
}


```

# **General checks**

## 1- Select multiple questions check

<p style="color:blue"> Checking if all select multiple questions/choices columns are present in the dataset </p>

```{r results='asis'}

sm_list <- questions %>% filter(q.type=="select_multiple") %>% select(name,list_name)

if(sum(!unlist(map2(sm_list[["name"]],sm_list[["list_name"]],
~paste0(.x,".",choices %>% filter(list_name == .y) %>% pull(name))
)) %in% colnames(data),na.rm = T) > 0) {
  colnames(data)[!unlist(map2(sm_list[["name"]],sm_list[["list_name"]],
~paste0(.x,".",choices %>% filter(list_name == .y) %>% pull(name))
)) %in% colnames(data)] %>% as.data.frame() %>% knitr::kable()

} else {
  colorize("Check passed ✓","green")
}
  
  
```

## 2- Unique UUIDs

<p style="color:blue"> Checking of duplicate surveys </p>

```{r results='asis'}

if (sum (duplicated(data$uuid),na.rm = T) > 0 ) {
  data[duplicated(data$uuid),] %>% head() %>% select(idp_code,uuid) %>% knitr::kable()
} else {
    colorize("Check passed ✓","green")
}

```

## 3- NAs per row

<p style="color:blue"> Number of NAs per row </p>

```{r}
hist(rowSums(is.na(data)),
  xlab = "Number of NA per row",
  main = "NAs per row",
  breaks = sqrt(nrow(data))
) 
```

<p style="color:blue"> I would recommend checking rows with more than 280 NAs </p>

```{r}

data$nb_na = rowSums(is.na(data))
data %>% filter(nb_na >= 280 ) %>% select(idp_code,uuid,nb_na) %>% head() %>% knitr::kable()

```

# **Numerical checks**

## 1- Outliers checks

<p style="color:blue"> Boxplots of numerical questions </p>

```{r echo=F}
numerical_questions <- questions %>% filter(type %in% c("integer","calculate")) %>% pull(name) %>% .[6:26] %>% .[!. %in% c("check_nb_all_selected","check_nb_none_selected")]

data_2 <- data %>% mutate_at(numerical_questions,as.numeric)
```


```{r, fig.show="hold", out.width="50%", echo=F}
par(mar = c(4, 4, 2, 2))
boxplot(data_2[[numerical_questions[1]]],
  main = numerical_questions[1]
)
boxplot(data_2[[numerical_questions[2]]],
  main = numerical_questions[2]
)
```


```{r, fig.show="hold", out.width="50%", echo=F}
par(mar = c(4, 4, 2, 2))
boxplot(data_2[[numerical_questions[3]]],
  main = numerical_questions[3]
)
boxplot(data_2[[numerical_questions[4]]],
  main = numerical_questions[4]
)
```


```{r, fig.show="hold", out.width="50%", echo=F}
par(mar = c(4, 4, 2, 2))
boxplot(data_2[[numerical_questions[5]]],
  main = numerical_questions[5]
)
boxplot(data_2[[numerical_questions[6]]],
  main = numerical_questions[6]
)
```


```{r, fig.show="hold", out.width="50%", echo=F}
par(mar = c(4, 4, 2, 2))
boxplot(data_2[[numerical_questions[7]]],
  main = numerical_questions[7]
)
boxplot(data_2[[numerical_questions[8]]],
  main = numerical_questions[8]
)
```


```{r, fig.show="hold", out.width="50%", echo=F}
par(mar = c(4, 4, 2, 2))
boxplot(data_2[[numerical_questions[9]]],
  main = numerical_questions[9]
)
boxplot(data_2[[numerical_questions[10]]],
  main = numerical_questions[10]
)
```


```{r, fig.show="hold", out.width="50%", echo=F}
par(mar = c(4, 4, 2, 2))
boxplot(data_2[[numerical_questions[11]]],
  main = numerical_questions[11]
)
boxplot(data_2[[numerical_questions[12]]],
  main = numerical_questions[12]
)
```

```{r, fig.show="hold", out.width="50%", echo=F}
par(mar = c(4, 4, 2, 2))
boxplot(data_2[[numerical_questions[13]]],
  main = numerical_questions[13]
)
boxplot(data_2[[numerical_questions[14]]],
  main = numerical_questions[14]
)
```


```{r, fig.show="hold", out.width="50%", echo=F}
par(mar = c(4, 4, 2, 2))
boxplot(data_2[[numerical_questions[15]]],
  main = numerical_questions[15]
)
boxplot(data_2[[numerical_questions[16]]],
  main = numerical_questions[16]
)
```


```{r, fig.show="hold", out.width="50%", echo=F}
par(mar = c(4, 4, 2, 2))
boxplot(data_2[[numerical_questions[17]]],
  main = numerical_questions[17]
)
boxplot(data_2[[numerical_questions[18]]],
  main = numerical_questions[18]
)
```

## 2- 999 values
<p style="color:blue"> 999 should be recoded to NA </p>

```{r results='asis'}
nb_999 <- ((data %>% mutate_all(as.character)) == "999") %>% as.data.frame() %>% summarise(across(.cols = everything(), sum, na.rm =T)) 
if (rowSums(nb_999)>0){
  nb_999 %>% select(names(nb_999)[nb_999[1,] > 0]) %>% t() %>% head() %>% knitr::kable()
} else {
    colorize("Check passed ✓","green")
}
```



# **DSA speceic checks**

## 1 - Ki_role_other 

<p style="color:blue"> Translate and recode options </p>

```{r results='asis'}
if (data %>% filter(!is.na(ki_role_other),ki_role=="other") %>% nrow()>0){
  data %>% filter(!is.na(ki_role_other),ki_role=="other") %>% select(ki_role,ki_role_other) %>% head() %>% knitr::kable()
} else {
  colorize("Check passed ✓","green")
}
```

## 2 - nb_shelters 
<p style="color:blue"> cccm_populationestimates_shelters < 15 </p>

```{r results='asis'}
if (data %>% filter(cccm_populationestimates_shelters<15) %>% nrow()>0){
  data %>% filter(cccm_populationestimates_shelters<15) %>% select(idp_code,cccm_populationestimates_shelters) %>% head() %>% knitr::kable()
} else {
  colorize("Check passed ✓","green")
}
```

## 3 - nb_families 
<p style="color:blue"> cccm_populationestimates_families < 15 </p>

```{r results='asis'}
if (data %>% filter(cccm_populationestimates_families<15) %>% nrow()>0){
  data %>% filter(cccm_populationestimates_families<15) %>% select(idp_code,cccm_populationestimates_families) %>% head() %>% knitr::kable()
} else {
  colorize("Check passed ✓","green")
}
```

## 4 - nb_individuals 
<p style="color:blue"> cccm_populationestimates_individuals < 15 </p>

```{r results='asis'}
if (data %>% filter(cccm_populationestimates_individuals<15) %>% nrow()>0){
  data %>% filter(cccm_populationestimates_individuals<15) %>% select(idp_code,cccm_populationestimates_individuals) %>% head() %>% knitr::kable()
} else {
  colorize("Check passed ✓","green")
}
```

## 5 - women_comm_rep
<p style="color:blue"> ki_role == women_comm_rep & women_comm_rep not selected in cccm_committees </p>

```{r results='asis'}
if (data %>% filter(ki_role == "women_comm_rep"
                & cccm_committees.women_committee==0) %>% nrow()>0){
  data %>% filter(ki_role == "women_comm_rep"
                & cccm_committees.women_committee==0) %>% select(idp_code) %>% head() %>% knitr::kable()
} else {
  colorize("Check passed ✓","green")
}
```



## 5 - evictions_landowner_other
<p style="color:blue"> Translate and recode options </p>

```{r results='asis'}
if (data %>% filter(!is.na(evictions_landowner_other),evictions_landowner=="other") %>% nrow()>0){
  data %>% filter(!is.na(evictions_landowner_other),evictions_landowner=="other") %>% select(evictions_landowner,evictions_landowner_other) %>% head() %>% knitr::kable()
} else {
  colorize("Check passed ✓","green")
}
```

## 6 - families and individuals
<p style="color:blue"> Checking number of familites against the number of individuals </p>


```{r results='asis'}
if (data %>% mutate(check_fam = cccm_populationestimates_individuals <= cccm_populationestimates_families)  %>%
  filter(check_fam == T) %>% nrow()>0){
  data %>% mutate(check_fam = cccm_populationestimates_individuals <= cccm_populationestimates_families)  %>%
  filter(check_fam == T) %>% select(idp_code,uuid) %>% head() %>% knitr::kable()
} else {
  colorize("Check passed ✓","green")
}
```


## 6 - Number of surveys / site
<p style="color:blue"> Checking high/low number of surveys collected per idp site </p>

```{r results='asis'}

data_3 <- data %>% group_by(idp_code) %>% summarise(n=n()) %>% filter(n<3 | n>6) %>% arrange(desc(n))

if (data_3 %>% nrow()>0){
  rbind(data_3 %>% head(5),data_3 %>% tail(5)) %>% select(idp_code,collected_surveys=n) %>% knitr::kable()
} else {
  colorize("Check passed ✓","green")
}
```

